A Parallel General Purpose Multi-Objective Optimization Framework, with Application To Electron Beam Dynamics

### 2019-03-19 paper/submission/prab.tex
resubmitted paper
edits to address reviewers second round of comments

Particle accelerators play a significant role in many aspects of science and
  technology.
Fields, such as material science, chemistry, the biosciences, particle
  physics, nuclear physics and medicine depend on reliable and effective
  particle accelerators, both as research and practical tools.
Achieving the required performance in the design, commissioning, and operation
  of accelerator facilities is a complex and versatile problem.
Despite the success of on-line models in some facilities~\cite{xiaobiao}, 
and various model dependent and model independent tuning and optimization techniques, 
empirical tuning by operators is a common method used at many research facilities.
When the beam dynamics is nonlinear, as is the case with space charge, 
simple and fast models are applicable only in a very restricted manner. 
This further complicates any mult-objective optimization by complicating the model. 
In order to be able to reliably identify optimal configurations of
  accelerators, we solve large multi-objective design optimization
  problems to automate the investigation for an optimal set of tuning parameters.
  This approach has been used in the past with much success~\cite{hofler13,bazarov05,yrss:09,jefferson,gull1,gull2,marija}.
The difference here being the implementation and application to a 
problem at the Argonne Wakefield Accelerator Facility (AWA).

@@

Several similar frameworks, e.g.~\cite{fide:09,lems:09,lbjt:07,dnld:06}, have
  been proposed.
Commonly these frameworks are tightly coupled to an optimization algorithm,
  e.g.\ only providing evolutionary algorithms as optimizers.
Users can specify optimization problems, but cannot change the
  optimization algorithm.
Our framework follows a more general approach, providing a user-friendly way
  to introduce new or choose from existing built-in multi-objective
  optimization algorithms.
Tailoring the optimization algorithm to the optimization problem at hand is
  an important feature due to the many different characteristics of
  optimization problems that should be handled by such a general framework.
As an example, it is shown how \textsc{Pisa}~\cite{pisa}, an existing evolutionary
  algorithm library, was integrated with ease.
Similarly, other multi-objective algorithms could be incorporated and
  used to solve optimization problems.

@@

% AA: this is too much for Phys Rev. AB

% \appendix

%\section{Forward Solver Implementation}
%
%A simple implementation, e.g.\ using a Python forward solver, is given below
%  in Listing 6.
%We assume the Python script executes a simulation and dumps results in the
%  \textsc{SDDS} file format~\cite{borl:00}.
%Utility classed are used to parse the result data and fill the appropriate
%  structures returned to the optimizer algorithm.
%
%\begin{code}[Simple Simulation Wrapper]
%#include <map>
%#include <string>
%#include <vector>
%
%#include "Util/Types.h"
%#include "Util/CmdArguments.h"
%#include "Simulation/Simulation.h"
%
%#include "boost/smart_ptr.hpp"
%
%class SimpleSimulation : public Simulation {
%
%public:
%
%    SimpleSimulation(Expressions::Named_t objectives,
%                     Expressions::Named_t constraints,
%                     Param_t params, std::string name, MPI_Comm comm,
%                     CmdArguments_t args);
%
%    ~SimpleSimulation();
%
%    /// Calls simulation through Python wrapper and returns when simulation
%    /// has either failed or finished.
%    void run() {
%        // user implements wrappers
%        prepare_input_file();
%        run_python_simulation();
%    }
%
%    /// Parse SDDS stat file and build up requested variable dictionary.
%    void collectResults(){
%
%        Expressions::Named_t::iterator it;
%        for(it = objectives_.begin(); it != objectives_.end(); it++) {
%
%            Expressions::Expr_t *objective = it->second;
%
%            // find out which variables we need in order to evaluate the
%            // objective
%            variableDictionary_t variable_dictionary;
%            std::set<std::string> req_vars = objective->getReqVars();
%
%            if(req_vars.size() != 0) {
%
%                boost::scoped_ptr<SDDSReader> sddsr(new SDDSReader(fn));
%
%                try {
%                    sddsr->parseFile();
%                } catch(OptPilotException &e) {
%                    std::cout << "Exception while parsing SDDS file: "
%                        << e.what() << std::endl;
%                    break;
%                }
%
%                // get all the required variable values from the stat file
%                foreach(std::string req_var, req_vars) {
%                    if(variable_dictionary.count(req_var) == 0) {
%
%                        try {
%                            double value = 0.0;
%                            sddsr->getValue(1 /*atTime*/, req_var, value);
%                            variable_dictionary.insert(
%                                std::pair<std::string, double>(req_var, value));
%                        } catch(OptPilotException &e) {
%                            std::cout << "Exception while getting value "
%                                      << "from SDDS file: " << e.what()
%                                      << std::endl;
%                        }
%                    }
%                }
%            }
%
%            // and evaluate the expression using the built dictionary of
%            // variable values
%            Expressions::Result_t result =
%                objective->evaluate(variable_dictionary);
%
%            std::vector<double> values;
%            values.push_back(boost::get<0>(result));
%            bool is_valid = boost::get<1>(result);
%
%            reqVarInfo_t tmps = {EVALUATE, values, is_valid};
%            requestedVars_.insert(
%                std::pair<std::string, reqVarInfo_t>(it->first, tmps));
%
%        }
%    }
%
%    /// returns container containing all requested variables with results
%    reqVarContainer_t getResults() { return requestedVars_; }
%
%private:
%
%    /// holds solutions returned to the optimizer
%    reqVarContainer_t requestedVars_;
%
%    Expressions::Named_t objectives_;
%    Expressions::Named_t constraints_;
%
%    MPI_Comm comm_;
%};
%
%#endif
%\end{code}

>>>>

Particle accelerators play a significant role in many aspects of science and
  technology.
Fields, such as material science, chemistry, the biosciences, particle
  physics, nuclear physics and medicine depend on reliable and effective
  particle accelerators, both as research and practical tools.
Achieving the required performance in the design, commissioning, and operation
  of accelerator facilities is a complex and versatile problem.
Despite the success of on-line models in some facilities~\cite{xiaobiao}, 
and various model dependent and model independent tuning and optimization techniques, 
empirical tuning by operators is a common method used at many research facilities.
When the beam dynamics is nonlinear, as is the case with space charge, 
simple and fast models are applicable only in a very restricted manner. 
This further complicates any mult-objective optimization by complicating the model. 
In order to be able to reliably identify optimal configurations of
  accelerators, we solve large multi-objective design optimization
  problems to automate the investigation for an optimal set of tuning parameters.
  This approach has been used in the past with much 
  success~\cite{bazarov05,yrss:09,chao09, hofler13,jefferson,gong15, gongthesis15,gull1,gull2,marija}.
The difference here being the implementation and application to a 
problem at the Argonne Wakefield Accelerator Facility (AWA).

@@

Several similar frameworks, e.g.~\cite{dnld:06, lbjt:07, fide:09,lems:09}, have
  been proposed.
Commonly these frameworks are tightly coupled to an optimization algorithm,
  e.g.\ only providing evolutionary algorithms as optimizers.
Users can specify optimization problems, but cannot change the
  optimization algorithm.
Our framework follows a more general approach, providing a user-friendly way
  to introduce new or choose from existing built-in multi-objective
  optimization algorithms.
Tailoring the optimization algorithm to the optimization problem at hand is
  an important feature due to the many different characteristics of
  optimization problems that should be handled by such a general framework.
As an example, it is shown how \textsc{Pisa}~\cite{pisa}, an existing evolutionary
  algorithm library, was integrated with ease.
Similarly, other multi-objective algorithms could be incorporated and
  used to solve optimization problems.

@@
